### 2017-04-04 paper/oopsla17-submission/evaluation-accuracy.tex

\paragraph{Our Classifiers}
We evaluate five classifiers, each trained on the full set of features.
% features: 44 local syntactic features, 176 contextual syntactic
% features, 55 typing features, and a single expression size feature.
% %
% \ES{should explain the make-up of these groups}
%
% We preemptively discard expressions that are not part of the minimal
% type error slice --- we will explain the rationale for this in
% \autoref{sec:feature-utility} --- and thus the final feature count is
% 276.
%
Our classifiers are:
%
\begin{enumerate}
\item A logistic regression trained with a learning rate
  $\alpha = 0.001$, an L2 regularization rate $\lambda = 0.001$, and a
  mini-batch size of 200.
\item A decision tree trained with the \ES{HUMA?} algorithm and \ES{HUMA: hyperparams?}.
\item A random forest trained with the \ES{HUMA?} algorithm and 30 estimators \ES{HUMA: hyperparams?}.
\item Two multi-layer perceptron neural networks, both trained with $\alpha = 0.001$,
  $\lambda = 0.001$, and a mini-batch size of 200. The first MLP
  contains a single hidden layer of 10 neurons, and the second contains
  a hidden layer of 500 neurons. This allows us to investigate how well
  the MLP can \emph{compress} its model (cf.~\cite{FIXME}).
\end{enumerate}
%
All classifiers were trained for 20 epochs on one dataset
--- \ie they were shown each program 20 times ---
before being evaluated on the other.

>>>>

\paragraph{Our Classifiers}
We evaluate five classifiers, each trained on the full set of features.
% features: 44 local syntactic features, 176 contextual syntactic
% features, 55 typing features, and a single expression size feature.
% %
% \ES{should explain the make-up of these groups}
%
% We preemptively discard expressions that are not part of the minimal
% type error slice --- we will explain the rationale for this in
% \autoref{sec:feature-utility} --- and thus the final feature count is
% 276.
%
Our classifiers are:
%
\begin{enumerate}
\item A logistic regression trained with a learning rate
  $\eta = 0.001$, an L2 regularization rate $\lambda = 0.001$, and a
  mini-batch size of 200.
\item A decision tree trained with the \ES{HUMA?} algorithm and \ES{HUMA: hyperparams?}.
\item A random forest trained with the \ES{HUMA?} algorithm and 30 estimators \ES{HUMA: hyperparams?}.
\item Two multi-layer perceptron neural networks, both trained with $\eta = 0.001$,
  $\lambda = 0.001$, and a mini-batch size of 200. The first MLP
  contains a single hidden layer of 10 neurons, and the second contains
  a hidden layer of 500 neurons. This allows us to investigate how well
  the MLP can \emph{compress} its model (cf.~\cite{FIXME}).
\end{enumerate}
%
All classifiers were trained for 20 epochs on one dataset
--- \ie they were shown each program 20 times ---
before being evaluated on the other.

### 2017-04-04 paper/oopsla17-submission/evaluation-utility.tex

We tested our hypothesis with the \linear and
%
\hiddenFH\footnote{A layer of 500 neurons is excessive when we have so few
  input features --- we use \hiddenFH for continuity with the
  surrounding sections.}
%
classifiers, cross-validated ($k=10$) over the combined SP14/FA15
dataset. We used a learning rate $\alpha=0.001$, L2 regularization rate
$\lambda=0.001$, and mini-batch size of 200. We trained for a single
epoch on feature sets (1) and (2), and for 8 epochs on (3), so that the
total number of training samples would be roughly equal for each feature
set.
%
\lstDeleteShortInline{|} % sigh...
In addition to the accuracy metric from before, we report each
classifier's \emph{recall} --- \ie ``How many true changes can we
remember?'' --- defined as
$$
\frac{|\mathsf{predicted} \cap \mathsf{oracle}|}
     {|\mathsf{oracle}|}
$$
where $\mathsf{predicted}$ is limited to the top 3 predictions of the
classier, and $\mathsf{oracle}$ is limited to the expressions that
changed \emph{and} were part of the type error slice. We make the
latter distinction as (1) changes that are not part of the type error
slice are noise in the data set, and (2) it makes the comparison between
feature sets easier to interpret since $\mathsf{oracle}$ never changes.
% NOTE: keep this at the end of the para or it screws up spacing...
\lstMakeShortInline{|}
%
\begin{table}[ht]
  \caption{
    Impact of Type Error Slice on Accuracy.
    \ES{TODO: load these numbers from CSV}
  }\label{tab:type-error-slice}
  \centering
  \begin{tabular}{lrcrrrrcrrrr}
    \toprule
                       &             & & \multicolumn{4}{c} \linear        & & \multicolumn{4}{c} \hiddenFH      \\
                                         \cmidrule{4-7}                        \cmidrule{9-12}
    Feature Set        & \# Features & & Top-1  & Top-2  & Top-3  & Recall & & Top-1  & Top-2  & Top-3  & Recall \\
    \midrule
    Local Syntax       & 44          & & 23.5\% & 42.5\% & 55.9\% & 36.5\% & & 28.9\% & 47.7\% & 58.6\% & 38.4\% \\
    + \InSlice         & 45          & & 46.1\% & 64.9\% & 75.4\% & 48.8\% & & 54.4\% & 70.8\% & 81.3\% & 56.2\% \\
    Filter by \InSlice & 44          & & 54.9\% & 71.4\% & 82.5\% & 57.6\% & & 56.6\% & 71.9\% & 82.5\% & 57.3\% \\
    \bottomrule
  \end{tabular}
\end{table}

@@

We investigate the relative impact of the other
three classes of features discussed in \autoref{sec:features}, assuming
we have discarded expressions not in the type error slice.
%
For this experiment we consider again a baseline of only local syntactic
features, extended by each combination of
%
(1) expression size;
(2) contextual syntactic features; and
(3) typing features.
%
As before, we perform a 10-fold cross-validation with $\alpha = 0.001$,
$\lambda = 0.001$, and a mini-batch size of 200, but we
train for a full 20 epochs to make the differences more apparent.
%
\begin{table}[ht]
  \caption{
    Impact of Contextual Features on Accuracy.
    \ES{TODO: load these numbers from CSV}
  }\label{tab:contextual-features}
  \centering
  \begin{tabular}{lrcrrrrcrrrr}
    \toprule
                             &             & & \multicolumn{4}{c} \linear        & & \multicolumn{4}{c} \hiddenFH      \\
                                               \cmidrule{4-7}                        \cmidrule{9-12}
    Feature Set              & \# Features & & Top-1  & Top-2  & Top-3  & Recall & & Top-1  & Top-2  & Top-3  & Recall \\
    \midrule
    Local Syntax             &  44         & & 55.0\% & 71.6\% & 82.6\% & 57.7\% & & 56.3\% & 72.1\% & 82.8\% & 57.6\% \\
    \midrule
    + Size                   &  45         & & 55.8\% & 72.8\% & 82.5\% & 57.3\% & & 60.5\% & 75.0\% & 83.8\% & 57.9\% \\
    + Context                & 220         & & 59.9\% & 77.7\% & 86.4\% & 63.0\% & & 71.4\% & 84.1\% & 90.8\% & 69.5\% \\
    + Types                  & 102         & & 62.2\% & 77.7\% & 85.7\% & 62.2\% & & 73.2\% & 84.8\% & 90.2\% & 69.2\% \\
    \midrule
    + Context + Size         & 221         & & 60.2\% & 77.9\% & 86.0\% & 62.5\% & & 71.7\% & 83.5\% & 90.7\% & 69.3\% \\
    + Types + Size           & 103         & & 62.0\% & 78.2\% & 85.6\% & 62.3\% & & 73.5\% & 85.8\% & 91.4\% & 70.5\% \\
    + Context + Types        & 275         & & 63.2\% & 80.3\% & 87.9\% & 65.4\% & & 77.3\% & 87.5\% & 92.4\% & 72.9\% \\
    \midrule
    + Context + Types + Size & 276         & & 62.3\% & 80.0\% & 88.0\% & 65.4\% & & 77.2\% & 87.9\% & 92.7\% & 72.9\% \\
    \bottomrule
  \end{tabular}
  % \begin{minipage}{0.49\linewidth}
  % \centering
  % \hiddenF
  % \begin{tabular}{lrrrr}
  %   \toprule
  %   Feature Set                 & Top-1  & Top-2  & Top-3  & Recall \\
  %   \midrule
  %   Local Syntax                & 56.9\% & 72.2\% & 82.8\% & 57.9\% \\
  %   \midrule
  %   + Size                      & 59.7\% & 74.6\% & 83.0\% & 57.4\% \\
  %   + Context                   & 70.9\% & 83.7\% & 90.4\% & 69.2\% \\
  %   + Types                     & 72.1\% & 84.1\% & 90.3\% & 69.3\% \\
  %   \midrule
  %   + Size + Context            & 69.8\% & 83.5\% & 90.2\% & 68.6\% \\
  %   + Size + Types              & 72.3\% & 84.6\% & 90.3\% & 69.5\% \\
  %   + Context + Types           & 75.5\% & 86.4\% & 91.5\% & 71.7\% \\
  %   \midrule
  %   + All                       & 75.0\% & 86.8\% & 91.9\% & 72.0\% \\
  %   \bottomrule
  % \end{tabular}
  % \end{minipage}
\end{table}
%
\paragraph{Results}
\autoref{tab:contextual-features} summarizes the results of this experiment.
%
The \linear classifier and the \hiddenFH start off
competitive when given only local syntactic features, but the \hiddenFH
quickly outperforms as more features are added.

>>>>

We tested our hypothesis with the \linear and
%
\hiddenFH\footnote{A layer of 500 neurons is excessive when we have so few
  input features --- we use \hiddenFH for continuity with the
  surrounding sections.}
%
classifiers, cross-validated ($k=10$) over the combined SP14/FA15
dataset. We used a learning rate $\eta=0.001$, L2 regularization rate
$\lambda=0.001$, and mini-batch size of 200. We trained for a single
epoch on feature sets (1) and (2), and for 8 epochs on (3), so that the
total number of training samples would be roughly equal for each feature
set.
%
\lstDeleteShortInline{|} % sigh...
In addition to the accuracy metric from before, we report each
classifier's \emph{recall} --- \ie ``How many true changes can we
remember?'' --- defined as
$$
\frac{|\mathsf{predicted} \cap \mathsf{oracle}|}
     {|\mathsf{oracle}|}
$$
where $\mathsf{predicted}$ is limited to the top 3 predictions of the
classier, and $\mathsf{oracle}$ is limited to the expressions that
changed \emph{and} were part of the type error slice. We make the
latter distinction as (1) changes that are not part of the type error
slice are noise in the data set, and (2) it makes the comparison between
feature sets easier to interpret since $\mathsf{oracle}$ never changes.
% NOTE: keep this at the end of the para or it screws up spacing...
\lstMakeShortInline{|}
%
\begin{table}[ht]
  \caption{
    Impact of Type Error Slice on Accuracy.
    \ES{TODO: load these numbers from CSV}
  }\label{tab:type-error-slice}
  \centering
  \begin{tabular}{lrcrrrrcrrrr}
    \toprule
                       &             & & \multicolumn{4}{c} \linear        & & \multicolumn{4}{c} \hiddenFH      \\
                                         \cmidrule{4-7}                        \cmidrule{9-12}
    Feature Set        & \# Features & & Top-1  & Top-2  & Top-3  & Recall & & Top-1  & Top-2  & Top-3  & Recall \\
    \midrule
    Local Syntax       & 44          & & 23.5\% & 42.5\% & 55.9\% & 36.5\% & & 28.9\% & 47.7\% & 58.6\% & 38.4\% \\
    + \InSlice         & 45          & & 46.1\% & 64.9\% & 75.4\% & 48.8\% & & 54.4\% & 70.8\% & 81.3\% & 56.2\% \\
    Filter by \InSlice & 44          & & 54.9\% & 71.4\% & 82.5\% & 57.6\% & & 56.6\% & 71.9\% & 82.5\% & 57.3\% \\
    \bottomrule
  \end{tabular}
\end{table}

@@

We investigate the relative impact of the other
three classes of features discussed in \autoref{sec:features}, assuming
we have discarded expressions not in the type error slice.
%
For this experiment we consider again a baseline of only local syntactic
features, extended by each combination of
%
(1) expression size;
(2) contextual syntactic features; and
(3) typing features.
%
As before, we perform a 10-fold cross-validation with $\eta = 0.001$,
$\lambda = 0.001$, and a mini-batch size of 200, but we
train for a full 20 epochs to make the differences more apparent.
%
\begin{table}[ht]
  \caption{
    Impact of Contextual Features on Accuracy.
    \ES{TODO: load these numbers from CSV}
  }\label{tab:contextual-features}
  \centering
  \begin{tabular}{lrcrrrrcrrrr}
    \toprule
                             &             & & \multicolumn{4}{c} \linear        & & \multicolumn{4}{c} \hiddenFH      \\
                                               \cmidrule{4-7}                        \cmidrule{9-12}
    Feature Set              & \# Features & & Top-1  & Top-2  & Top-3  & Recall & & Top-1  & Top-2  & Top-3  & Recall \\
    \midrule
    Local Syntax             &  44         & & 55.0\% & 71.6\% & 82.6\% & 57.7\% & & 56.3\% & 72.1\% & 82.8\% & 57.6\% \\
    \midrule
    + Size                   &  45         & & 55.8\% & 72.8\% & 82.5\% & 57.3\% & & 60.5\% & 75.0\% & 83.8\% & 57.9\% \\
    + Context                & 220         & & 59.9\% & 77.7\% & 86.4\% & 63.0\% & & 71.4\% & 84.1\% & 90.8\% & 69.5\% \\
    + Types                  & 102         & & 62.2\% & 77.7\% & 85.7\% & 62.2\% & & 73.2\% & 84.8\% & 90.2\% & 69.2\% \\
    \midrule
    + Context + Size         & 221         & & 60.2\% & 77.9\% & 86.0\% & 62.5\% & & 71.7\% & 83.5\% & 90.7\% & 69.3\% \\
    + Types + Size           & 103         & & 62.0\% & 78.2\% & 85.6\% & 62.3\% & & 73.5\% & 85.8\% & 91.4\% & 70.5\% \\
    + Context + Types        & 275         & & 63.2\% & 80.3\% & 87.9\% & 65.4\% & & 77.3\% & 87.5\% & 92.4\% & 72.9\% \\
    \midrule
    + Context + Types + Size & 276         & & 62.3\% & 80.0\% & 88.0\% & 65.4\% & & 77.2\% & 87.9\% & 92.7\% & 72.9\% \\
    \bottomrule
  \end{tabular}
  % \begin{minipage}{0.49\linewidth}
  % \centering
  % \hiddenF
  % \begin{tabular}{lrrrr}
  %   \toprule
  %   Feature Set                 & Top-1  & Top-2  & Top-3  & Recall \\
  %   \midrule
  %   Local Syntax                & 56.9\% & 72.2\% & 82.8\% & 57.9\% \\
  %   \midrule
  %   + Size                      & 59.7\% & 74.6\% & 83.0\% & 57.4\% \\
  %   + Context                   & 70.9\% & 83.7\% & 90.4\% & 69.2\% \\
  %   + Types                     & 72.1\% & 84.1\% & 90.3\% & 69.3\% \\
  %   \midrule
  %   + Size + Context            & 69.8\% & 83.5\% & 90.2\% & 68.6\% \\
  %   + Size + Types              & 72.3\% & 84.6\% & 90.3\% & 69.5\% \\
  %   + Context + Types           & 75.5\% & 86.4\% & 91.5\% & 71.7\% \\
  %   \midrule
  %   + All                       & 75.0\% & 86.8\% & 91.9\% & 72.0\% \\
  %   \bottomrule
  % \end{tabular}
  % \end{minipage}
\end{table}
%
\paragraph{Results}
\autoref{tab:contextual-features} summarizes the results of this experiment.
%
The \linear classifier and the \hiddenFH start off
competitive when given only local syntactic features, but the \hiddenFH
quickly outperforms as more features are added.

### 2017-04-04 paper/oopsla17-submission/introduction.tex
Wes pass over Section 2.
address some of wes' fixmes

Type systems are one of the most widely-used techniques for program
verification. A type checker only accepts a program if it can
\emph{prove} the absence of an entire class of bugs.
%
Typed functional languages like \ocaml and \haskell are equipped with
particularly powerful type checkers that can operate without a single
annotation, alleviating a prime concern with verification techniques.
%
Yet this power comes with a cost: since type annotations signify the
programmer's \emph{intent}, without them the checker must
make assumptions about what the programmer intended to do.
%
This results in the common phenomenon of type errors being reported far
from their source, an annoyance to veterans of the type system, but a
major hurdle for novices~\cite{FIXME,FIXME}.

\begin{figure}[ht]
\begin{minipage}{0.45\linewidth}
\begin{ecode}
  let rec sumList xs =
    match xs with
    | []   -> __[]__
    | h::t -> h __+__ ==sumList t==
\end{ecode}
\end{minipage}
\begin{minipage}{0.49\linewidth}
\begin{verbatim}
File "sumList.ml", line 4, characters 16-25:
  This expression has type 'a list
  but an expression was expected of type int
\end{verbatim}
\end{minipage}
\caption{(left) An ill-typed \ocaml program that should sum the elements of a
  list. The \underline{underlined} expressions are equally valid
  locations to blame. The expression blamed by the \ocaml compiler
  is in \textbf{bold}.

  FIXME: This bolding is ambiguous, because ``let rec'', ``match'' and
  ``with'' are also bolded (by \\ecode)! You need to find another way to
  highlight what ocaml is yelling about.

  (right) The error reported by \ocaml.}
\label{fig:sumList}
\end{figure}
%
Consider the |sumList| program in \autoref{fig:sumList}, written by a
student in an undergraduate Programming Languages course.
%
The program is meant to sum the integers in a list, but the student has
accidentally given the empty list as the base case, rather than 0.
%
The \ocaml compiler collects typing constraints as it traverses the
program, and reports an error the moment it finds an inconsistent
constraint.
%
In this case it blames the recursive call to |sumList|, complaining that
|sumList| returns a |list| where an |int| was expected by the |+|
operator. Note that this blame is inconsistent with the programmer's
intention and thus may not be helpful debugging information to a novice.

>>>>

Type systems are one of the most widely-used techniques for program
verification. A type checker only accepts a program if it can
\emph{prove} the absence of an entire class of bugs.
%
Typed functional languages like \ocaml and \haskell are equipped with
particularly powerful type checkers that can operate without a single
annotation, alleviating a prime concern with verification techniques.
%
Yet this power comes with a cost: since type annotations signify the
programmer's \emph{intent}, without them the checker must
make assumptions about what the programmer intended to do.
%
This results in the common phenomenon of type errors being reported far
from their source, an annoyance to veterans of the type system, but a
major hurdle for novices \citep{Wand1986-nw,Joosten1993-yx}.

\begin{figure}[ht]
\begin{minipage}{0.45\linewidth}
\begin{ecode}
  let rec sumList xs =
    match xs with
    | []   -> __[]__
    | h::t -> h __+__ ==sumList t==
\end{ecode}
\end{minipage}
\begin{minipage}{0.49\linewidth}
\begin{verbatim}
File "sumList.ml", line 4, characters 16-25:
  This expression has type 'a list
  but an expression was expected of type int
\end{verbatim}
\end{minipage}
\caption{(left) An ill-typed \ocaml program that should sum the elements of a
  list. The \underline{underlined} expressions are equally valid
  locations to blame. The expression blamed by the \ocaml compiler
  is in \textbf{bold}.
  %
  % FIXME: This bolding is ambiguous, because ``let rec'', ``match'' and
  % ``with'' are also bolded (by \\ecode)! You need to find another way to
  % highlight what ocaml is yelling about.
  %
  (right) The error reported by \ocaml.}
\label{fig:sumList}
\end{figure}
%
Consider the |sumList| program in \autoref{fig:sumList}, written by a
student in an undergraduate Programming Languages course.
%
The program is meant to sum the integers in a list, but the student has
accidentally given the empty list as the base case, rather than 0.
%
The \ocaml compiler collects typing constraints as it traverses the
program, and reports an error the moment it finds an inconsistent
constraint.
%
In this case it blames the recursive call to |sumList|, complaining that
|sumList| returns a |list| where an |int| was expected by the |+|
operator. Note that this blame is inconsistent with the programmer's
intention and thus may not be helpful debugging information to a novice.
